---
title: "Alpha Diversity Analysis"
author: "Brittany Seibert"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

# Before you begin:

These scripts were tailored for the analyses performed in: 

Seibert et al, 2022, *Pathobiology and dysbiosis of the respiratory and intestinal microbiota in 14 months old Golden Syrian hamsters infected with SARS-CoV-2*

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Purpose: 
The purpose of this script is to analyze alpha and beta diversity of the different groups within the lungs. This will correspond to figures 

# Load the needed packages
```{r,message=FALSE}
library(vegan)
library(phyloseq)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(microbiome)
library(harrietr)
library(ggforce)
```

# Load the R file from previous data processing
```{r,message=FALSE}
# Load the file RarifiedASVs.rds 
Phylo.samples_l_filter <- readRDS(file = "/Users/Phylo.samples_l.rds")
```

## Before performing alpha diveristy analysis, the sequences will be rarified. 
Rarify the data using the vegan package 

### Before rarifying the data in vegan, convert the phyloseq object to a dataframe that is worke-able in vegan
```{r}
# Create a function to convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# Create a function to convert the count table within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}

# Use the function on the lung data 
Phylo.samples_l_filter
# Use the function to create a dataframe of the sample data 
sampledataveg_l<- pssd2veg(Phylo.samples_l_filter)
saveRDS(sampledataveg_l, "sampledataveg_l.rds")

# Use the function to create a matrix of the ASV table. This will be a matrix that has ASV number as column headers and sample IDs as row names 
ASVtableveg_l <- psotu2veg(Phylo.samples_l_filter)
```

### Rarify the SAMPLES in vegan
```{r}
# Calculate total OTU counts for all samples (sum all columns)
total.counts <- apply(ASVtableveg_l, 1, sum) 

# List all sample sums in descending order - useful for deciding what your sample cutoff should be
sort(total.counts, decreasing=TRUE)

# List the minimum coverage (sequences per sample)
min(total.counts)

# Change as appropriate for samples. We will include all samples
sample.size_l <- 13306 

# Subset sample.shared to remove samples with < sample.size sequences
sample.used_l <- ASVtableveg_l[total.counts>=sample.size_l,] 

# Resample all samples to a total of sample.size sequences (I do not want to rarify again so i will load sample.rare from rarified data previously analyzed)
#sample.rare_l <- rrarefy(sample.used_l, sample=sample.size_l) 

# Save sample.rare to a file so that i can reload the rarified file 
#saveRDS(object = sample.rare_l, file = "/Users/baseibert/Perez_Lab/Projects/Microbiome/Projects/Hamsters_SARS/R_Files/R_Files/RarifiedASVs_l.rds")

# Load the file RarifiedASVs.rds 
sample.rare_l <- readRDS(file = "/Users/baseibert/Perez_Lab/Projects/Microbiome/Projects/Hamsters_SARS/R_Files/R_Files/RarifiedASVs_l.rds")

# Visualize the rarefaction curve
library("phyloseq.extended")
rare_dpc_lung <- ggrare(Phylo.samples_l_filter, step = 500, color = "dpi")

pdf("/Users/baseibert/Perez_Lab/Projects/Microbiome/Projects/Hamsters_SARS/R_Files/OutputGraphs/Lungs/RarefactionCurve_lungs_dpc.pdf",
    width = 4, height = 3) 

rare_dpc_lung + 
  scale_fill_gradientn(breaks = c(3,6), 
                      colours = c("deeppink", "limegreen"))+
  scale_color_gradientn(breaks = c(3,6), 
                      colours = c("deeppink", "limegreen"))+
  geom_vline(xintercept = 13306, color = "black", linetype="dashed", size = 0.6)+
  scale_y_continuous(breaks = seq(0, 400, by=100), limits=c(0, 400))+
  scale_x_continuous(breaks = seq(0, 80000, by=20000), limits=c(0, 90000))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  theme(axis.title.x=element_text(colour = "black"),
          axis.title.y = element_text(colour = "black"),
          axis.text.x = element_text(colour = "black"),
          axis.text.y = element_text(colour = "black"),
          axis.line.x = element_line(color="black"),
          axis.line.y = element_line(color="black"),
          legend.text = element_text(color="black"))+
  theme(axis.line = element_line(color="black", size = 0.2),
          panel.border = element_blank())

dev.off()

```

**Alpha diversity**
<style>
div.rose { background-color:#ffc4ef; border-radius: 10px; padding: 40px;}
</style>
<div class = "rose">
Estimating alpha diversity of microbial communities is problematic no matter what you do. The best is to subsample the libraries with replacement to estimate the species abundance of the real population while standardizing sampling effort.
</div>
<br>

### Import the rarified data into phyloseq since that is how we are going to calculate the alpha diversity metrics 
```{r}
# Import metadata file 
metadata <- read.csv("/Users/Sars_hamster_metadata.csv", header=T, row.names=1, check.names=F)

# Import rarified count into phyloseq
psrare_l <- otu_table(sample.rare_l, taxa_are_rows = FALSE)

# Import taxonomy file 
taxa_l <- as.matrix(read.table("/Users/ASVs_taxonomy_dada_wSpecies_lungs.tsv", header=T, row.names=1, check.names=F, sep="\t"))

# Import taxonomy table into phyloseq 
tax_l = tax_table(taxa_l)

# Import the sample metadata into phyloseq object
sample = sample_data(metadata)

# Merge the count table, taxonomy table and sample data 
Phylo.rare_l  = merge_phyloseq(psrare_l, tax_l, sample)
Phylo.rare_l

# Save sample.rare to a file so that i can reload the rarified file 
saveRDS(object = Phylo.rare_l, file = "/Users/Phylo.rare_l.rds")
```

### Calculate the alpha diversity metrics and plot the graphs in ggplot 
```{r}
# Import metadata file 2 with sampled ID as a . instead of a -
metadata_2 <- read.csv("/Users/Sars_hamster_metadata_2.csv", header=T, row.names=1, check.names=F)

# Calculate diversity metrics in phyloseq 
PhlyoRichness_l <- estimate_richness(Phylo.rare_l, measures = c("Observed", "Shannon", "InvSimpson", "Fisher"))
# Make a column that contains the rownames so that we have a column of the SampleIDs
PhlyoRichness_l$SampleID <- rownames(PhlyoRichness_l)

PhlyoRichness_l_evenness <- evenness(Phylo.rare_l, index = "all", zeroes = TRUE, detection = 0)
# Make a column that contains the rownames so that we have a column of the SampleIDs
PhlyoRichness_l_evenness$SampleID <- rownames(PhlyoRichness_l_evenness)
PhlyoRichness_l_evenness$SampleID <- gsub("-",".",PhlyoRichness_l_evenness$SampleID)

PhlyoRichness_l <- merge(PhlyoRichness_l, PhlyoRichness_l_evenness, by = "SampleID")

# Join the diversity tables with metadata 
PhlyoRichness_meta_l = left_join(PhlyoRichness_l, metadata_2, by = "SampleID")

# Export data 
write.csv(PhlyoRichness_meta_l, "/Users/AlphaDiversityIndexes_lungs.csv")

# Set the colors for all of the graphs 
colorgroups = c("coral2", "dodgerblue1", "white")
```

### Figure 3A (left): Visualize the number of observed ASVs in the different groups in the lungs
```{r}
# Filter so that you only have shannon values
PhlyoRichness_meta_l.observed <- PhlyoRichness_meta_l[,c(2,16)]

# Re-level the groups 
PhlyoRichness_meta_l.observed$Group_PostInfect<- factor(PhlyoRichness_meta_l.observed$Group_PostInfect, levels = c('SARS-CoV-2', 'IAV-SARS-CoV-2', 'Mock'))

# FIGURE 3A
ggplot(PhlyoRichness_meta_l.observed, aes(x=Group_PostInfect, y=Observed, fill = Group_PostInfect))+ 
  geom_boxplot()+
  geom_point(shape = 21, 
             size = 3,
             colour = "Black",
             show.legend = F,
             alpha = 0.3, 
             position = position_dodge(0.75))+
  scale_fill_manual(values = colorgroups)+
  scale_y_continuous(breaks = seq(0, 400, by=100), limits=c(0, 400))+
  ylab("Observed")+
  theme_gray()+
  guides(fill=guide_legend(title = "Group", override.aes=list(shape=21)))+
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(vjust=1.5, size = 17),
        axis.text.x = element_text(size = 16, colour = "black"),
        axis.text.y = element_text(size = 18, colour = "black"),
        strip.text.y = element_text(angle = 0),
        legend.position="none")+
  theme(panel.grid = element_blank(),
          axis.line = element_line(color="black", size = 0.5),
          panel.border = element_blank(),
          panel.background = element_blank())
  stat_compare_means(method = "kruskal.test")
  
# Filter for only duodenum and then perform pairwise comparison 
compare_means(Observed ~ Group_PostInfect,  data = PhlyoRichness_meta_l.observed, method = "wilcox.test", p.adj = "bonferroni")
```

### Figure 3A (right): Visualize the number of shannon diversity in the different groups in the lungs
```{r}
# Filter so that you only have shannon values
PhlyoRichness_meta_l.shannon <- PhlyoRichness_meta_l[,c(3,16)]

# Re-level the groups 
PhlyoRichness_meta_l.shannon$Group_PostInfect<- factor(PhlyoRichness_meta_l.shannon$Group_PostInfect, levels = c('SARS-CoV-2', 'IAV-SARS-CoV-2', 'Mock'))

# Set the colors for all of the graphs 
colorgroups = c("coral2", "dodgerblue1", "white")

# FIGURE 3A
ggplot(PhlyoRichness_meta_l.shannon, aes(x=Group_PostInfect, y=Shannon, fill = Group_PostInfect))+ 
  geom_boxplot()+
  geom_point(shape = 21, 
             size = 3,
             colour = "Black",
             show.legend = F,
             alpha = 0.3, 
             position = position_dodge(0.75))+
  scale_fill_manual(values = colorgroups)+
  scale_y_continuous(breaks = seq(0, 6, by=2), limits=c(0, 6))+
  ylab("Shannon")+
  theme_gray()+
  guides(fill=guide_legend(title = "Group", override.aes=list(shape=21)))+
  theme(axis.title.x=element_blank(),
        axis.title.y = element_text(vjust=1.5, size = 17),
        axis.text.x = element_text(size = 16, colour = "black"),
        axis.text.y = element_text(size = 18, colour = "black"),
        strip.text.y = element_text(angle = 0),
        legend.position="none")+
  theme(panel.grid = element_blank(),
          axis.line = element_line(color="black", size = 0.5),
          panel.border = element_blank(),
          panel.background = element_blank())
  stat_compare_means(method = "kruskal.test")
  
# Statistics testing for comparison among means of the bray curtis for different groups (this uses the Wilcoxon Sign test (non-parametric))

# Filter for only duodenum and then perform pairwise comparison 
compare_means(Shannon ~ Group_PostInfect,  data = PhlyoRichness_meta_l.shannon, method = "wilcox.test", p.adj = "bonferroni")
```

# Beta Diversity 

## Figure 3B AND S3A: Visualize Jaccard (3B) Bray-Curtis (S3A) distances of the lungs
```{r}
# FIGURE 3B
# Calculating relative abundance and creating new dataframe with relative abundance data
ASVtableveg_l.rel <-         
  decostand(sample.rare_l, method = "total")

# Calculate the Jaccard distances 
# The Wisconsin transformation normalizes 0-1 so it will be very similar to the unweighted so i will not use the transformation 
lung_jaccard_nmds <-  metaMDS(ASVtableveg_l.rel, 
        distance="jaccard",         # set the diversity metric used
        binary = TRUE,  #https://stats.stackexchange.com/questions/242110/nmds-from-jaccard-and-bray-curtis-identical-is-that-a-bad-thing
        k=3,                     # set number of dimensions
        maxit = 999, 
        trymax=500, 
        autotransform=FALSE)

# Determine the stress value on the number of dimensions
lung_jaccard_nmds

# Create a stress plot to visualize the data 
stressplot(lung_jaccard_nmds)

# Plot the bray-curtis distances in simple plot 
ordiplot(lung_jaccard_nmds, 
         disp="sites")
orditorp(lung_jaccard_nmds, "sites")

metadata_edited <- metadata %>%
  filter(SampleType == "Lung")

MyMeta = data.frame(
  sites = c("BS-592", "BS-593", "BS-594", "BS-595", "BS-596", "BS-597", "BS-598", "BS-599", "BS-600", "BS-601", "BS-602", "BS-587", "BS-588", "BS-589", "BS-590", "BS-591"),
  Groups = c("SARS2", "SARS2","SARS2", "SARS2","IAV-SARS2", "IAV-SARS2", "IAV-SARS2", "IAV-SARS2", "IAV-SARS2", "IAV-SARS2", "IAV-SARS2", "Mock", "Mock", "Mock", "SARS2", "SARS2"),
  row.names = "sites")

MyMeta <- c("SARS2", "SARS2","SARS2", "SARS2","FLUAV-SARS2", "FLUAV-SARS2", "FLUAV-SARS2", "FLUAV-SARS2", "FLUAV-SARS2", "FLUAV-SARS2", "FLUAV-SARS2", "Mock", "Mock", "Mock", "SARS2", "SARS2")
colors <- c("coral2", "coral2", "coral2", "coral2", "dodgerblue1", "dodgerblue1", "dodgerblue1", "dodgerblue1", "dodgerblue1", "dodgerblue1", "dodgerblue1", "black", "black", "black", "coral2", "coral2")
shape <- c(15, 19, 19, 19, 15, 15, 15, 15, 19, 19, 19, 15, 15, 19, 15, 15) 

plot(lung_jaccard_nmds, 
         disp="sites")
ordiellipse(lung_jaccard_nmds, MyMeta, kind = "sd", label = T, col = c("dodgerblue1", "black", "coral2"))
points(lung_jaccard_nmds, disp="sites", pch=shape, col = colors, cex = 1)

# FIGURE S3A
# Calculate the bray-curtis distances 
# The Wisconsin transformation normalizes 0-1 so it will be very similar to the unweighted so i will not use the transformation 
lung_bray_nmds<-  metaMDS(ASVtableveg_l.rel, 
          distance="bray",         # set the diversity metric used
          k=3,                     # set number of dimensions
          maxit = 999, 
          trymax=500, 
          autotransform=FALSE)

# Determine the stress value on the number of dimensions
lung_bray_nmds

# Create a stress plot to visualize the data 
stressplot(lung_bray_nmds)

# Plot the bray-curtis distances in simple plot 
ordiplot(lung_bray_nmds, 
         disp="sites")
orditorp(lung_bray_nmds, "sites")

plot(lung_bray_nmds, 
         disp="sites")
ordiellipse(lung_bray_nmds, MyMeta, kind = "sd", label = T, col = c("dodgerblue1", "black", "coral2"))
points(lung_bray_nmds, disp="sites", pch=shape, col = colors, cex = 1)
```

### Perform PERMANOVA analysis to statistically analyze if there are differences among the different groups within the lungs
```{r}
# PERMANOVA ANALYSIS 
Phylo.samples_i_lung <- subset_samples(Phylo.rare_l)
pseq.rel <- microbiome::transform(Phylo.samples_i_lung, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)
```

### Figure 3C: Calculate the jaccard distances and produce a boxplot for the lungs
```{r}
# Calculating relative abundance and creating new dataframe with relative abundance data
ASVtableveg_l.rel <-         
  decostand(sample.rare_l, method = "total")

# Calculate the jaccard distances of the lungs using vegan 
jaccard_veg_lung<- vegdist(ASVtableveg_l.rel, 
                          method = "jaccard", 
                          binary = TRUE)

# Turn the distance class into a matrix  
jaccard_veg_lung_matrix <- as.matrix(jaccard_veg_lung)

# Use program harrietr to melt the distance matrix into 3 column data frame
jaccard_veg_lung_df <- melt_dist(jaccard_veg_lung_matrix)

# Convert the Var 1 and Var2 to a character 
jaccard_veg_lung.melt2 = jaccard_veg_lung_df %>%
  filter(as.character(iso1) != as.character(iso2)) %>%
  mutate_if(is.factor, as.character)

# Only keep columns SampleID and GroupGen
sd1 = metadata %>%
  dplyr::select("SampleID", "Group_PostInfect") %>%
  mutate_if(is.factor,as.character)

# Convert the column names to Var1 and Type 1 in metadata sheet sd
colnames(sd1) = c("iso1", "Group1")

# Join the distance matrix and the metadata sheet sd
jaccard_veg_lung.melt2.sd = left_join(jaccard_veg_lung.melt2, sd1, by = "iso1")

# Convert the column names to Var2 and Type 2 in metadata sheet sd
colnames(sd1) = c("iso2", "Group2")

# Join the distance matrix and the metadata sheet sd
jaccard_veg_lung.melt2.sd = left_join(jaccard_veg_lung.melt2.sd, sd1, by = "iso2")

write.csv(jaccard_veg_lung.melt2.sd, "/Users/jaccard_lung.csv")

# The spreadsheet was edited to include 2 columns: group1 and then the group 2. The matrix was converted into a dataframe. 
jaccard.melt2.sd_lung_edited <- read.csv("/Users/jaccard_lung_edited_2.csv", 
                     header=T, 
                     row.names=1, 
                     check.names=F)

# Reorder the groups so that they are in the same order
jaccard.melt2.sd_lung_edited$Name<- factor(jaccard.melt2.sd_lung_edited$Name, levels =c('SARS2', 'FLUAV-SARS2', 'Mock', 'SARS2 v. Mock', 'FLUAV-SARS2 v. Mock', 'SARS2 v. FLUAV-SARS2'))

colorgroups = c( "coral2", "dodgerblue1", "black", "darkgoldenrod3", "darkgoldenrod3", "darkgoldenrod3")

# FIGURE 3C
pdf("/Users/Jaccard_Boxplot_Lung_Group.pdf",
    width = 11, height = 8.4) # Width and height in inches 2.5 x 3 for single digits and 2.7 x 3 for double digit legends

ggplot(jaccard.melt2.sd_lung_edited, aes(x = Name, y = dist, color = Name)) + 
    geom_boxplot() +
    scale_color_manual(values = colorgroups)+
    #scale_color_identity() +
    scale_y_continuous(breaks = seq(0.2, 1, by=.2), limits=c(0.1, 1.1))+
    ylab("Jaccard Distance")+
    guides(fill=guide_legend(title = "Group", override.aes=list(shape=21)))+
    #theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(vjust=1.5, size = 22),
          axis.text.x = element_text(angle=40, hjust=1, colour = "black", size = 22),
          axis.text.y = element_text(colour = "black", size = 26),
          axis.line.x = element_line(color="black", size = 0.5),
          axis.line.y = element_line(color="black", size = 0.5),
          legend.position = "none")+
    theme(panel.grid = element_blank(),
            axis.line = element_line(color="black", size = 0.5),
            panel.border = element_blank(),
            panel.background =  element_rect(fill='white'))

dev.off()

stats <- compare_means(dist ~ Name,  data = jaccard.melt2.sd_lung_edited, method = "wilcox.test", p.adj = "bonferroni")  
stats

write.csv(stats, "/Users/Lungs_JaccardComparison.csv")
```

### Figure S3B: Then, calculate the bray-curtis distances and produce a boxplot for the lung
```{r}
# Calculating relative abundance and creating new dataframe with relative abundance data
ASVtableveg_l.rel <-         
  decostand(sample.rare_l, method = "total")

# Calculate the bray-curtis distances of the duodenum using vegan 
bray_veg_lung <- vegdist(ASVtableveg_l.rel, 
                          method = "bray", 
                          binary = FALSE)

# Turn the distance class into a matrix  
bray_veg_lung_matrix <- as.matrix(bray_veg_lung)

# Use program harrietr to melt the distance matrix into 3 column data frame
bray_veg_lung_df <- melt_dist(bray_veg_lung_matrix)

# Convert the Var 1 and Var2 to a character 
bray_veg_lung.melt2 = bray_veg_lung_df %>%
  filter(as.character(iso1) != as.character(iso2)) %>%
  mutate_if(is.factor, as.character)

# Only keep columns SampleID and GroupGen
sd1 = metadata %>%
  dplyr::select("SampleID", "Group_PostInfect") %>%
  mutate_if(is.factor,as.character)

# Convert the column names to Var1 and Type 1 in metadata sheet sd
colnames(sd1) = c("iso1", "Group1")

# Join the distance matrix and the metadata sheet sd
bray_veg_lung.melt2.sd = left_join(bray_veg_lung.melt2, sd1, by = "iso1")

# Convert the column names to Var2 and Type 2 in metadata sheet sd
colnames(sd1) = c("iso2", "Group2")

# Join the distance matrix and the metadata sheet sd
bray_veg_lung.melt2.sd = left_join(bray_veg_lung.melt2.sd, sd1, by = "iso2")

write.csv(bray_veg_lung.melt2.sd, "/Users/bray_lung.csv")

# The spreadsheet was edited to include 2 columns: group1 and then the group 2. The matrix was converted into a dataframe. 
bray.melt2.sd_lung_edited <- read.csv("/Users/bray_lung_edited.csv", 
                     header=T, 
                     row.names=1, 
                     check.names=F)

# Reorder the groups so that they are in the same order
bray.melt2.sd_lung_edited$Name<- factor(bray.melt2.sd_lung_edited$Name, levels =c('SARS2', 'FLUAV-SARS2', 'Mock', 'SARS2 v. Mock', 'FLUAV-SARS2 v. Mock', 'SARS2 v. FLUAV-SARS2'))

colorgroups = c( "coral2", "dodgerblue1", "black", "darkgoldenrod3", "darkgoldenrod3", "darkgoldenrod3")

# FIGURE S3B
pdf("/Users/BrayCurtis_Boxplot_Lung_Group.pdf",
    width = 11, height = 8.4) # Width and height in inches 2.5 x 3 for single digits and 2.7 x 3 for double digit legends

ggplot(bray.melt2.sd_lung_edited, aes(x = Name, y = dist, color = Name)) + 
    geom_boxplot() +
    scale_color_manual(values = colorgroups)+
    scale_y_continuous(breaks = seq(0.2, 1, by=.2), limits=c(0.1, 1.1))+
    ylab("Bray-Curtis Distance")+
    guides(fill=guide_legend(title = "Group", override.aes=list(shape=21)))+
    #theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(vjust=1.5, size = 22),
          axis.text.x = element_text(angle=40, hjust=1, colour = "black"),
          axis.text.y = element_text(colour = "black"),
          text = element_text(size=25),
          axis.line.x = element_line(color="black", size = 0.5),
          axis.line.y = element_line(color="black", size = 0.5),
          legend.position = "none")+
    theme(panel.grid = element_blank(),
            axis.line = element_line(color="black", size = 0.5),
            panel.border = element_blank(),
            panel.background =  element_rect(fill='white'))

dev.off()

stats <- compare_means(dist ~ Name,  data = bray.melt2.sd_lung_edited, method = "wilcox.test", p.adj = "bonferroni")    
stats
write.csv(stats, "/Users/Lungs_BrayComparison.csv")
```
