---
title: "Sample Processing for Lung Hamster Samples"
output: html_notebook
---

# Before you begin:

These scripts were tailored for the analyses performed in:
  
Seibert et al, 2022, *Pathobiology and dysbiosis of the respiratory and intestinal microbiota in 14 months old Golden Syrian hamsters infected with SARS-CoV-2.*

# Purpose: 
This file will go through the initial sample processing and filtering to remove phyla that do not have an assigned Phyla, any Eukaryotes, chloroplasts and mitochondria, 

# Load the needed packages
```{r,message=FALSE}
library(phyloseq)
library(dplyr)
library(ggplot2)
library(metagMisc)
```

## Import files from dada2 scripts 

First, we will need to import the ASV count table, taxonomy file and the metadata file.\ 

*In order to read the metadata into a phyloseq object you need to make the SampleID (first column) the row headers*
```{r}
# Import metadata file (provided)
metadata <- read.csv("/path/Sars_hamster_metadata.csv", header=T, row.names=1, check.names=F)

# Import count table 
count_l <- read.table("/path/ASVs_counts.tsv", header=T, row.names=1, check.names=F, sep="\t")

# Import taxonomy file 
taxa_l <- as.matrix(read.table("/path/ASVs_taxonomy_dada_wSpecies.tsv", header=T, row.names=1, check.names=F, sep="\t"))
```

## Filtering variants for ONLY BACTERIA  

When the organisms from a set of samples are well-represented in the taxonomic reference database, it is advisable to filter variants in which a high-rank taxonomy could not be assigned. These are most likely sequence artifacts that don’t exist in nature. Therefore, we will filter out any variants that are not classified as bacteria, have an unclassified phyla or are identified as Chloroplast or Mitochondria.\

### Import files into PhyloSeq
```{r}
# Import count table into phyloseq
ps_l = otu_table(count_l, taxa_are_rows = TRUE)

# Import taxonomy table into phyloseq 
tax_l = tax_table(taxa_l)

# Import metadata into phyloseq object
sample = sample_data(metadata)

# Merge the count table, taxonomy table and sample data 
Phylo_l = merge_phyloseq(ps_l, tax_l, sample)
Phylo_l
```

### First we will filter out any Phyla that are unassigned
```{r}
# Create table, number of features for each phyla
table(tax_table(Phylo_l)[, "Phylum"], exclude = NULL)

# Remove the ASVs and name as uncharacterized if the Phylum has an NA 
Phylo.filter_l <- subset_taxa(Phylo_l, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized"))
table(tax_table(Phylo.filter_l)[, "Phylum"], exclude = NULL)
```

### Next we will filter out Eukaryotes, chloroplasts and mitochondria, because we only intended to amplify bacterial sequences
```{r}
Phylo.filter_l # Print information about the number of taxa and samples before filtering 
table(tax_table(Phylo.filter_l)[, "Kingdom"], exclude = NULL) # Print a table that states the number of sequences for the different kingdoms before filtering

Phylo.filter_filter_l <- Phylo.filter_l %>%
  subset_taxa(
    Kingdom == "Bacteria" &
    Class   != "Chloroplast" &
    Family  != "Mitochondria")

Phylo.filter_filter_l # Print information about the number of taxa and samples after filtering 
table(tax_table(Phylo.filter_filter_l)[, "Kingdom"], exclude = NULL) # Print a table that states the number of sequences for the different kingdoms after filtering
```

### Now we will separate the biological samples from the negative sequencing controls 
```{r}
# Subset the samples using column sampleClass in metadata
Phylo.samples_l <- Phylo.filter_filter_l %>%
  subset_samples(sampleClass == "sample")
Phylo.samples_l
```


### Next we will filter OTUs that are at low prevalence to try and remove sequencing artificacts. 

We will filter out any OTUs found in at least 5% of samples and total OTU abundance is greater that 10 reads 
```{r}
# Filter the samples with the specified conditions 
Phylo.samples_l_filter # Print information about the number of taxa and samples before filtering 
Phylo.samples_l_filter <- phyloseq_filter_prevalence(Phylo.samples_l, prev.trh = 0.05, abund.trh = 10, threshold_condition = "AND")
Phylo.samples_l_filter # Print information about the number of taxa and samples after filtering 
```

# Save the filtered phyloseq objects 
```{r}
# Save phyloseq files to a R file so that i can reload the filtered file 
saveRDS(object = Phylo.samples_l_filter, file = "/path/Phylo.samples_l.rds")
```
